Garaphlet Analysis for HiC data

### 2018-03-09 papers/presentation_2018_03_08.tex
improved

            \end{itemize}
        \end{column}
        \begin{column}{.5\textwidth}
             \begin{figure}[H]
                \centering
                \includegraphics[width=\textwidth]{figures/grand_scheme_of_things.png}
                \caption*{}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\frame{
    \frametitle{HiC Data: An Overal Picture}
     \begin{figure}[H]
        \centering
        \includegraphics[width=.8\textwidth]{figures/grand_scheme_of_things.png}
        \caption*{}
    \end{figure}

}

\frame {
    \begin{figure}
        \centering
        \includegraphics[width=.8\textwidth, height=.82\textheight]{figures/normal-b-cell-all-heatmaps.jpeg}
        \caption*{\cite{wang2013properties}}
    \end{figure}

}

\section{Strategies}
\begin{frame}
    \frametitle{Graphlets}
    \framesubtitle{Definitions}
    Graphlet comparison, introduced by \citeA{prvzulj2007biological},
    is a novel method used to compare large networks in order to
    find local similarities in them.
    \begin{itemize}
        \item \textbf{Fragment:} A connected subgraph.
        \item \textbf{Motifs:} Fragments that occur with a frequency much higher than
            that occuring in a randomly generated graph.
        \item \textbf{Graphlets:} An arbitrary, induced fragment.
        An edge is the only two-node graphlet.
        \item \textbf{Induced graphs:} Given a graph $G(V, E)$ and $S \subseteq V$,
            then $G'(S, E')$
            is a graphlet iff $E' = \{(u, v) | u, v \in V \text{ and } 
            (u, v) \in E \rightarrow (u, v) \in E'\}$
        \item \textbf{Orbits:} Set of all nodes in a graphlet that can be
            swapped with each other while not changing the graph.
    \end{itemize}
\end{frame}

@@

    \textit{Important Result}: Proteins with similar surroundings perform
    similar functions.
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/signature-comparison.png}
        \caption*{\cite{milenkoviae2008uncovering}}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Introduction}
    \textbf{\citeA{milenkovic2010cancer}}:

@@

        \item Predict the cancer-relatedness of a protein $i$ using
            an enrichment criteria $\frac{k}{|C_i|}$ \\
            \begin{itemize}
                \item $C_i$ : the cluster where protein $i$ belongs\\
                \item $k$ : the number of cancer-causing proteins in $C_i$ \\
                \item $|C_i|$ : the size of $C_i$
            \end{itemize}
    \end{enumerate}
\end{frame}

\frame {
    \frametitle{Stage I: Thresholding}
    For thresholding, I use \textit{local thresholding} methods.
    \begin{enumerate}
        \item Create a zero matrix with the same size as the contact 
            map (M)
        \item Slide a kernel through each pixel
        \item If the pixel $(i, j)$  satisfies a particular condition
            the set $A[i, j]$.
    \end{enumerate}
    \bigskip
    Conditions that I considered:
    \begin{enumerate}
        \item If the pixel is local maximum with respect to the pixels
            that fall in the kernel \\
            $A_{ij} > A_{i',j'} 
            \hspace{2cm}  (i', j') \in \{i-k, i+k\}\times \{j-k, j+k\}$
        \item If the pixel value is larger that some standard deviation
            from the mean of the values that fall inside the kernel.
            $A_{ij} > mean(A_{i',j'}) + std(A_{i',j'})$ \\
    \end{enumerate}
}

\frame {
    \frametitle{Kernels}
    \begin{figure}[H]
        \centering
        \includegraphics[width=.9\textwidth]{figures/kernels.png}
        \caption*{}
    \end{figure}
}

\frame {
    \frametitle { Extracting graphlets }
    \begin{enumerate}
        \item I used the \texttt{orca} library in \texttt{R} programming 
        language.
        \item For each loci, a \textit{signature vector} of length 73 is 
        created.
        \item For each chromosome, an $N_c \times 73$, matrix is returned.
        where $N_c$ is the number of loci in chromosome $c$.
    \end{enumerate}
}

\frame {
    \frametitle { Finding distance between loci }
    In order to see how different two loci are in 
    terms of local neighborhood, we need to compare
    their corresponding \textit{signature values}.
    The distance measure proposed by \citeA{prvzulj2007biological}:

    \begin{equation}
        d_i = w_i \times \frac{log(S_i+1) - log(S'_i+1)}{log(max(S_i, S'_i) + 2)}
    \end{equation}
    The distance between $S$ and $S'$ can be calculated as follows:
    \begin{equation}
        D = \sum_{0}^{72}{d_i^2}
    \end{equation}
}

\frame {
    \frametitle{loci-loci distances between cell lines}
    I calculated loci-loci distance between cell lines using the formula
    above:
    \begin{figure}[H]
        \centering
        \includegraphics[width=.6\textwidth]{figures/loci_wise_distance.png}
        \caption*{\small {$A_{ij}$ in matrices above denotes distance between loci
        $i$ in row cell line and loci $j$ in the column cell line.} }
    \end{figure}
}

\frame {
    \frametitle{Loci-loci distances}
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/cross_chromosomal_signature_distance.png}
        \caption*{}
    \end{figure}
}

\frame {
    \frametitle{Loci-loci distances}
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/cross_chromosomal_signature_distance_separated.png}
        \caption*{cross-chromosomal signature distance with 1 standard deviation error bars}
    \end{figure}
}

\frame {
    \frametitle{Orbit correlations}
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/orbit_wise_correlation_chr1.png}
        \caption*{\small {$A_{ij}$ in matrices above denotes correlation between 
        orbit $i$ in row cell line and orbit $j$ in the column cell line.} }
    \end{figure}
}

\begin{frame}[allowframebreaks]
    \frametitle{References}
    \bibliographystyle{apacite}
    \bibliography{lit}
\end{frame}

>>>>

            \end{itemize}
        \end{column}
        \begin{column}{.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=.9\textwidth]{figures/normal-b-cell-all-heatmaps.jpeg}
                \caption*{\cite{wang2013properties}}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}
\section{Strategies}
\begin{frame}
    \frametitle{Graphlets}
    \framesubtitle{Definitions}
    Graphlet comparison, introduced by \citeA{prvzulj2007biological},
    is a novel method used to compare large networks in order to
    find local similarities in them.
    \begin{itemize}
        \item \textbf{Fragment:} A connected subgraph.
        \item \textbf{Motifs:} Fragments that occur with a frequency much higher than
            that occuring in a randomly generated graph.
        \item \textbf{Graphlets:} An arbitrary, induced fragment.
        An edge is the only two-node graphlet.
        \item \textbf{Induced graphs:} Given a graph $G(V, E)$ and $S \subseteq V$,
            then $G'(S, E')$
            is a graphlet iff $E' = \{(u, v) | u, v \in V \text{ and } 
            (u, v) \in E \rightarrow (u, v) \in E'\}$
        \item \textbf{Orbits:} Set of all nodes in a graphlet that can be
            swapped with each other while not changing the graph.
    \end{itemize}
\end{frame}

@@

    \textit{Important Result}: Proteins with similar surroundings perform
    similar functions.
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/signature-comparison.png}
        \caption*{\cite{milenkoviae2008uncovering}}
    \end{figure}
\end{frame}
\begin{frame}
    \frametitle{Introduction}
    \textbf{\citeA{milenkovic2010cancer}}:

@@

        \item Predict the cancer-relatedness of a protein $i$ using
            an enrichment criteria $\frac{k}{|C_i|}$ \\
            \begin{itemize}
                \item $C_i$ : the cluster where protein $i$ belongs\\
                \item $k$ : the number of cancer-causing proteins in $C_i$ \\
                \item $|C_i|$ : the size of $C_i$
            \end{itemize}
    \end{enumerate}
\end{frame}
\begin{frame}
    \frametitle{Graphlets}
    \framesubtitle{Challenges}
    \begin{enumerate}
        \item HiC contact maps are noisy. How do we de-noise them?
        \item Current applications of graphlets were on unweighted
            graphs, while HiC contact maps are weighted.
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{Balanced Network Deconvolution}
    \framesubtitle{Introduction}
    Proposed by \citeA{feizi2013network}, Balanced Network Deconvolution, is 
    a method that can be used to remove \textit{indirect effects} from a graph.
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/nd.png}
        \caption*{ \cite{feizi2013network}}
    \end{figure}
\end{frame}
\begin{frame}
    \frametitle{Balanced Network Deconvolution}
    \framesubtitle{Details}
    They assume that:
    \begin{equation}
        \label{indirect_effects}
        G_{obs} = G_{dir} + G_{dir}^2 + G_{dir} + ...
    \end{equation}
    They assume also that both $G_{obs}$ and $G_{dir}$ can be 
    eigen-decomposed and they have the same eigen-vectors:
    \begin{equation}
        G_{dir} = X \Sigma_{dir} X^T
    \end{equation}
    \begin{equation}
        G_{obs} = X \Sigma_{obs} X^T
    \end{equation}
    \begin{equation}
        G_{obs} = X \Sigma_{obs} X^T = X ( \Sigma_{dir} + \Sigma_{dir}^2 + ...) X^T
    \end{equation}
\end{frame}
\begin{frame}
    \frametitle{Balanced Network Deconvolution}
    \framesubtitle{Details}
    They also assume that eigen-values of the direct network are all
    between -1 and 1, i.e. 
    \begin{equation}
        -1 < \lambda^{dir}_i < 1 \quad \forall 1 \le i \le n
    \end{equation}
    \begin{equation}
        \Sigma_{obs} = \Sigma_{dir} + \Sigma_{dir}^2 + ...
    \end{equation}
    \begin{equation}
        \lambda_i^{obs} = \sum_{j=1}^{\infty} \lambda^{dir}_{ij} \quad\quad \forall i = 1 ... n
    \end{equation}
    \begin{equation}
        \lambda_i^{obs} = \frac{\lambda_i^{dir}}{1 - \lambda_i^{dir}}
    \end{equation}
    \begin{equation}
       \lambda_i^{dir} = \frac{\lambda_i^{obs}}{1 + \lambda_i^{obs}} 
    \end{equation}
\end{frame}
\begin{frame}
    \frametitle{Statistical Parametric Network (SPN)}
    Developed by \citeA{ginestet2011statistical}.
    \begin{figure}
        \centering
        \includegraphics[width=.5\textwidth]{figures/thresholded_adjacency.png}
        \caption*{An study of neuron connectivity
            networks among 43 subjects for 4 different memory tasks 
            (0-back, 1-back, 2-back and 3-back), and the
            resulting thresholded networks \cite{ginestet2011statistical}}
        \label{fig:thresholded_adjacency}
    \end{figure}
\end{frame}
\section{Results Thus Far}
\begin{frame}
    \frametitle{Histogram of frequencies in HiC}
    \begin{figure}[H]
        \centering
        \begin{subfigure}[b]{.45\textwidth} 
            \includegraphics[width=\textwidth]{figures/hist01.png}
            \caption{Interval $[0, 0.1]$}
            \label{fig:hist01}
        \end{subfigure}
        \begin{subfigure}[b]{.45\textwidth}
            \includegraphics[width=\textwidth]{figures/hist02.png}
            \caption{Interval $[0.1, 1]$}
            \label{fig:hist02}
        \end{subfigure}
        \caption{\ref{fig:hist01} is a histogram of intensities from
        chromosome 1 
        As can be seen, the two histograms are orders of magnitude
        different in terms of frequency.}
        \label{fig:hists}
    \end{figure}
\end{frame}
\begin{frame}
    \frametitle{Histogram of logarithm}
    \begin{figure}[H]
        \centering
        \includegraphics[width=1.05\textwidth]{figures/pyr_hist_1mb.png}
        \caption*{}
        \label{fig:pyr_hist_1mb}
    \end{figure}
\end{frame}
\begin{frame}
    \frametitle{Different Normalization methods}
    \begin{figure}[H]
        \centering
        \begin{subfigure}[b]{.25\textwidth}
            \includegraphics[width=\textwidth]{figures/original_cleaned.png}
            \caption{}
            \label{fig:original_cleaned}
        \end{subfigure}
        \begin{subfigure}[b]{.25\textwidth}
            \includegraphics[width=\textwidth]{figures/pearson.png}
            \caption{}
            \label{fig:pearson}
        \end{subfigure}
        \begin{subfigure}[b]{.25\textwidth}
            \includegraphics[width=\textwidth]{figures/scn.png}
            \caption{}
            \label{fig:scn}
        \end{subfigure}
        \begin{subfigure}[b]{.25\textwidth}
            \includegraphics[width=\textwidth]{figures/scn+pearson.png}
            \caption{}
            \label{fig:scn+pearson}
        \end{subfigure}
        \begin{subfigure}[b]{.25\textwidth}
            \includegraphics[width=\textwidth]{figures/pearson+scn.png}
            \caption{}
            \label{fig:pearson+scn}
        \end{subfigure}
        \begin{subfigure}[b]{.25\textwidth}
            \includegraphics[width=\textwidth]{figures/pyr.png}
            \caption{}
            \label{fig:pyr}
        \end{subfigure}
        \caption*{}
        \label{fig:normalizations}
    \end{figure}
\end{frame}
\begin{frame}[allowframebreaks]
    \frametitle{References}
    \bibliographystyle{apacite}
    \bibliography{lit}
\end{frame}
